\huge Tor's Usability for Censorship Circumvention

### 2016-04-12 pets2017-paper/pets2017-paper.tex
serge minor feedback changes.

\section{Related Work} 
There have been been no published usability evaluations of
Tor Browser since the 4.0 series, which introduced radical UI changes. 
The most recent usability effort is an unpublished pilot study  by Lee and Fifield~\cite {uxsprint} 
that tests the download, install, and  user interface of Tor Browser.  This study uncovered a number of issues~\cite{uxsprint2015-tickets}. Changes made are reflected in series 5.1 and later.

There have been three published user studies on Tor. Clark et. al examined various deployment
options for Tor Browser, such as Vidalia, Privoxy, Torbutton, and Foxyproxy, and found that none of them 
were satisfactory from a usability perspective ~\cite{clark2007usability}. Fabian et. al show that Tor's added latency 
~\cite{dingledine2009performance} causes users
to be frustrated, cancel requests more often, and prevents user adoption ~\cite{fabian2010privately}. 
Norcie et. al found found that 
64\% of users are unable to continue with installation or browsing at least once due to difficulties ~\cite{norcie2012eliminating}.

@@

This puts the burden of knowing technical information and configuring correctly on the user. Note that bridges, pluggable transports, and proxies are technical concepts that are likely unfamiliar to the average user. Although burdensome, the interface takes the approach of guiding the user through this technical process in order to give users their own agency in configuring their connection. No automation means that the user can have control over when there will be network traces of their activity trying to connect to Tor, and being able to connect successfully to the Tor the first time, provided they have all of the relevant information.

@@

\section{Goals/Evaluation Criteria}
\label{sec:goals}
We evaluate the existing and redesigned Tor Browser configuration interface 
through the empirical evaluation and heuristic evaluation criteria below. The
empirical evaluation criteria were considered during our qualitative and 
quantitative user studies, whereas the heuristic evaluation was considered during
the redesign process.

A heuristic evaluation is a rule-based evaluation that leverages standards
that have shown to work in the User Interface design 
community. We draw from Jacob Nielsen's ten heuristics~\cite{nielsen1994heuristic}
and other industry standards to form the following heuristic evaluation criteria: \\

@@

\subsubsection{Setup}
\label{sec:setup}
We simulated three censorship environments for our experiment.
These reproducible, stable environments are informed by our experience 
with pluggable transports and knowledge of commonly seen censorship 
techniques. Their goal in our user study is not to replicate the network 
environment in any particular country, but to require our distinct configurations
so that interface paths are tested. \\

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Mild censorship (E1)}: 
(Representative of countries such as France and Australia.)
Certain domains are blocked. Reaching these 
domains requires a censorship circumvention 
tool. The default option to ``connect'' to the Tor network 
directly will circumvent this censor. Additional correct
bridge or proxy configurations are optional.

@@

\item {\bfseries Comprehensive censorship (E3)}:
(Representative of countries such as China and Syria.)
Certain domains are blocked. Censorship circumvention tools
are thoroughly blocked. Tor is blocked by blocking all public
Tor relay nodes, and the censor has examined source code to block
all hard-coded bridge relays in the configuration interface. The default option
to ``connect'' to the Tor network directly will fail. Most bridges will fail,
but ``meek-amazon,'' ``meek-azure,'' and ``meek-google'' still work.
This is because domain-fronting requires censors to block entire CDNs to also
block this transport (which will cause huge collateral blocking damage), making
it resistant to aggressive censorship environments.
(See ~\cite{fifield2015blocking} for additional details.)\\
\end{itemize}

@@

\subsubsection{Procedure}
We conducted a qualitative evaluation of the interface by creating an environment 
for users to interact with the configuration interface in a censored environment, 
observing their interactions in real time without interacting with the participants, 
and following up with interviews about their experiences.

The one-hour, single-participant procedure begins when a participant enters a small 
room with a single computer, which is equipped with Tor, Chrome, Firefox, Internet Explorer, 
Chrome, and VLC (for screen recording). A participant is firstly informed of 
the risks of the study and consenting to data collection. If they consent, the 
experiment begins. A researcher informs the participant that they are in a
simulated censorship environment, where some websites and services are blocked. We
instruct them to visit a sample blocked website and a sample non-blocked website on a
non-Tor browser of their choice to illustrate the situation.
For the full script of what participants have heard, see Appendix~\ref{qualitative-script}.

After illustrating the censorship environment, participants are asked to 
complete a worksheet that asks to visit one blocked website and one non-blocked website. 
We chose Wikipedia's featured article of the day as the blocked website and 
the CNN homepage as our non-blocked website because the familiarity 
that most users have with these websites makes the browsing task relatively easy, 
which focuses participants' attention to configuring Tor Browser.

After instructions, researchers stepped out of the room so that there was no interaction
between the participant and researcher for the rest of the session. Participants' screens 
were recorded and streamed to another room, where the researchers were able to 
observe how a participant configured their browser. Participants had an 
average of 45 minutes to complete their worksheet. 
At this point, the participants do not know the details of their censorship environment,
only that they are actively being censored. Ultimately, participants needed to configure Tor Browser to 
circumvent the simulated censorship.

@@

\subsection{Results} 
In this section, we discuss problems encountered by our participants at each step of the configuration process and solutions to those problems. We conclude with our goals for improving the configuration interface. Because of the small sample size of qualitative studies, we can only state that these are problems that participants can encounter when configuring. All quotes are paraphrased from live transcriptions of participant interviews.\\

@@

\subsubsection{Bridge Configuration} 
When a participant chooses to configure their connection, their next step is to determine whether they need to configure a bridge, and to configure one if necessary. Answering yes to ``Does your Internet Service Provider (ISP) block or otherwise censor connections to the Tor Network'' directs to a bridge configuration screen, whereas answering no skips that screen. We observed that multiple participants struggled with the technical language on the screen, and ultimately decided to try process of elimination. Most participants with the default option to not configure a bridge.

@@

\section{Redesigning the Configuration Interface}
%skip talking about the process of redesigning the interface. It isn't that interesting.

% Heuristic evaluation metrics, from above: 
% \begin{enumerate} \itemsep1pt \parskip0pt \parsep0pt
%     \item {\bfseries Visibility of System Status} provide feedback on what the system is doing.
%     \item {\bfseries User Control and Freedom} allow users to override system, supports undo and redo actions. 
%     \item {\bfseries Error Management} expressing errors in plain language and offering a solution. 
%     \item {\bfseries Understandable Language} ~6th grade reading level. 
%     \item {\bfseries Minimal Design} suppressing rarely used information to reduce cognitive load. 
% \end{enumerate}

\begin{enumerate}
\item {\bfseries Major Problem}: Error messages do not express errors in plain language nor offer solutions. 
\item {\bfseries Major Problem}: The text was too technical for an average user to understand. 
\item {\bfseries Minor Problem}: The configuration is not visible to the users before they connect or while they are connecting.
\item {\bfseries Cosmetic Problem}: The interface can be simplified visually.  
 \end{enumerate}

\begin{figure*}[t]
	\centering
		\includegraphics[width=1.0\textwidth]{new-flow.pdf} 
		\caption{new flow} 
\end{figure*}

@@

@@

We also made additional changes to unburden the user (5-8) and build a mental model (9-10):\\

\begin{enumerate}
\setcounter{enumi}{4}
\item {\bfseries Added guidance on choosing connect vs configure.} We labeled the configure option as advanced, manual, and only for heavily censored situations. 
\item {\bfseries Got rid of proxy and bridge questions.}  We removed the bridge and proxy questions, which were highly technical and challenging for users to answer. There are two fewer screens in the interface.
\item {\bfseries Added auto-detect for proxies.} Auto-detection of proxies can be done locally, by scanning system files. This is safe even for at-risk users, and greatly simplifies the configuration process. 
\item {\bfseries Added explicit advice on choosing bridges.} The default bridge is still obfs3. There is added text that advises users to try a meek bridge if obfs3 does not work.

@@

We instrumented Tor Launcher 5.0.3 and our redesigned Tor Launcher.
The instrumentation logs recorded every meaningful interaction with the interface---every
button press, every menu selection, every screen change---along with a timestamp.
The logs enabled us to calculate statistics such as the success rate
for each condition.

We recorded the computer screen throughout each session.
Review of the videos revealed details not captured by the instrumentation logs alone,
such as participants' web searching and inspection of system networking settings.

@@

Out of our {\color {red} TODO} participants, ages ranged from {\color {red} TODO}
(sigma = {\color {red} TODO}). {\color {red} TODO} of our participants had at least
a college education.

>>>>

The qualitative user study gives insight into possible issues with
the deployed interface. 
The redesign process was guided by the participant observations, 
heuristic evaluations and empirical goals. 
Quantitative user studies provide empirical evaluations of the deployed
and redesigned interface.

\section{Related Work} 
There have been three published user studies on Tor. Clark et. al examined various deployment
options for Tor Browser, such as Vidalia, Privoxy, Torbutton, and Foxyproxy, and found that none of them 
were satisfactory from a usability perspective ~\cite{clark2007usability}. Fabian et. al show that Tor's added latency 
~\cite{dingledine2009performance} causes users
to be frustrated, cancel requests more often, and prevents user adoption ~\cite{fabian2010privately}. 
Norcie et. al found found that 
64\% of users are unable to continue with installation or browsing at least once due to difficulties ~\cite{norcie2012eliminating}.

There have been been no published usability evaluations of
Tor Browser since the 4.0 series, which introduced radical UI changes. 
The most recent usability effort is an unpublished pilot study  by Lee and Fifield~\cite {uxsprint} 
that tests the download, install, and  user interface of Tor Browser.  This study uncovered a number of issues~\cite{uxsprint2015-tickets}. Changes made are reflected in series 5.1 and later.

@@

This puts the burden of knowing technical information and configuring correctly on the user. Note that bridges, pluggable transports, and proxies are technical concepts that are likely unfamiliar to the average user. Although burdensome, the interface takes the approach of guiding the user through this technical process in order to give users their own agency in configuring their connection. There has been no previous work to explore the option of automation, but a manual configuration was preferred by Tor developers since it allows the user can have control over the process. If a user had all the relevant information, they would be able to connect without obvious network traces of attempting to connect to Tor.

@@

\section{Goals/Evaluation Criteria}
\label{sec:goals}
We evaluate the existing and redesigned Tor Browser configuration interface 
through the empirical evaluation and heuristic evaluation criteria below.

A heuristic evaluation is a rule-based evaluation based on time-tested standards
in the user interface design community. We draw from Jacob Nielsen's ten heuristics~\cite{nielsen1994heuristic}
and other industry standards to form the following heuristic evaluation criteria: \\

@@

\subsubsection{Setup}
\label{sec:setup}
We simulated three censorship environments for our qualitative and quantitative user studies. \\

\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item {\bfseries Mild censorship (E1)}: 
(Representative of countries such as France and Australia.)
Certain domains are blocked. Reaching these 
domains requires a censorship circumvention 
tool like Tor, but Tor is not explicitly blocked. 
The default option to ``connect'' to the Tor network 
directly will circumvent this censor. Additional correct
bridge or proxy configurations are optional.

@@

\item {\bfseries Comprehensive censorship (E3)}:
(Representative of countries such as China and Syria.)
Certain domains are blocked. Censorship circumvention tools
are thoroughly blocked. Tor is blocked by blocking all public
Tor relay nodes, and the censor has examined source code to block
all hard-coded bridge relays in the configuration interface. The default option
to ``connect'' to the Tor network directly will fail. Most bridges will fail,
but ``meek-amazon,'' ``meek-azure,'' and ``meek-google'' still work.
This is because domain-fronting requires censors to block entire CDNs to also
block this transport, which causes huge collateral blocking damage that
makes the censorship very visible and inconvenient. Currently, even
aggressive censorship environments resistant blocking entire CDNs.
(See ~\cite{fifield2015blocking} for additional details.)
\end{itemize}

These reproducible, stable environments are informed by our experience 
with pluggable transports and knowledge of commonly seen censorship 
techniques. Their goal in our user study is not to replicate the network 
environment in any particular country, but to require our distinct configurations
so that interface paths are tested.

@@

\subsubsection{Procedure}
We conducted a qualitative evaluation of the interface through real-time observations of 
users interacting with the configuration interface in a simulated censored environment.
There was no researcher-user interaction during the experiment. A follow-up interview 
detailed their thoughts about the configuration process.

The one-hour, single-participant procedure begins when a participant enters a small 
room with a single computer, which is equipped with Tor, Chrome, Firefox, Internet Explorer, 
and VLC (for screen recording). A participant is firstly informed of 
the risks of the study and consenting to data collection. If they consent, the 
experiment begins. A researcher reads a script (Appendix~\ref{qualitative-script}) that 
informs the participant of the simulated censorship environment and
instruct them to visit a sample blocked website and a sample non-blocked website on a
non-Tor browser of their choice to illustrate the situation. Then, participants are asked to 
complete a worksheet (Appendix ~\ref{participant-worksheet}) that gives information
on their censorship environment and instructs them
to visit one blocked website and one non-blocked website.

We chose Wikipedia's featured article of the day as the blocked website and 
the CNN homepage as our non-blocked website because the familiarity 
that most users have with these websites makes the browsing task relatively easy, 
which focuses participants' attention to configuring Tor Browser. 
After instructions, researchers stepped out of the room so that there was no interaction
between the participant and researcher for the rest of the session. Participants' screens 
were recorded and streamed to another room, where the researchers were able to 
observe how a participant configured their browser. Participants had an 
average of 45 minutes to complete their worksheet.

At this point, the participants do not know the details of their censorship environment,
only that they are actively being censored. Ultimately, participants needed to configure Tor Browser to 
circumvent the simulated censorship.

@@

\subsection{Results} 
In this section, we discuss problems encountered by our participants at each step of the configuration process and solutions to those problems. We conclude with our goals for improving the configuration interface. Because of the small sample size of qualitative studies, we can only state that these are problems that participants can encounter when configuring. All quotes are not verbatim, but from best-effort live transcriptions of participant interviews.

@@

\subsubsection{Bridge Configuration} 
When a participant chooses to configure their connection, their next step is to determine whether they need to configure a bridge, and to configure one if necessary. Answering yes to ``Does your Internet Service Provider (ISP) block or otherwise censor connections to the Tor Network'' directs to a bridge configuration screen, whereas answering no skips that screen. We observed that multiple participants struggled with the technical language on the screen, and ultimately decided to try process of elimination. Most participants went with the default option to not configure a bridge.

@@

\section{Redesigning the Configuration Interface}
\begin{figure*}[t]
	\centering
		\includegraphics[width=1.0\textwidth]{new-flow.pdf} 
		\caption{new flow} 
\end{figure*}

@@

\begin{itemize}
\item {\bfseries Major Problem}: Error messages do not express errors in plain language nor offer solutions. 
\item {\bfseries Major Problem}: The text was too technical for an average user to understand. 
\item {\bfseries Minor Problem}: The configuration is not visible to the users before they connect or while they are connecting.
\item {\bfseries Cosmetic Problem}: The interface can be simplified visually.  
 \end{itemize}

%HEURISTICS
%\begin{enumerate}
%    \item  {\bfseries Visibility of system status}: provide feedback on what the system is doing. 
%    \item  {\bfseries Error management}: expressing errors in plain language and offering a solution. 
%    \item  {\bfseries Understandable language}: explaining the task, rather than technical concepts. 
%    \item  {\bfseries Simple design}: showing options on an as-needed basis. 
%\end{enumerate}

@@

%EMPIRICALS
%\begin{enumerate}
%    \item {\bfseries Task Completion}: Almost all can successfully connect to Tor
%    \item {\bfseries  Time to Completion}: Time to completion. 
%    \item {\bfseries Safe for High-Risk Users}: It should be possible to configure Tor with the interface that it doesn't leak that they are using Tor. 
%\end{enumerate}

Other changes improve empirical goals from Section~\ref{sec:goals}: task completion, time to completion, and safe for high-risk users. 5-6 improve task completion, 7-8 reduce time to completion, and 9-10 builds users' mental models of censorship circumvention and network components involved. Users' full control of the configuration process remains preserved.\\

\begin{enumerate}
\setcounter{enumi}{4}
\item {\bfseries Added guidance on choosing connect vs configure.} We labeled the configure option as advanced, manual, and only for heavily censored situations. 
\item {\bfseries Added explicit advice on choosing bridge transports.} The default bridge is still obfs3. There is added text that advises users to try a meek bridge if obfs3 does not work.

\item {\bfseries Eliminated technical questions.}  We removed questions that determined whether a bridge and proxy should be configured, which were highly technical and challenging for users to answer. There are two fewer screens in the interface.
\item {\bfseries Added auto-detect for proxies.} Auto-detection of proxies can be done locally, by scanning system files. This is safe even for at-risk users, and greatly simplifies the configuration process.

@@

We instrumented Tor Launcher 5.0.3 and our redesigned Tor Launcher.
The instrumentation logs recorded every meaningful interaction with the interface---every
button press, every menu selection, every screen change---along with a timestamp.
The logs enabled us to calculate statistics such as the success rate
for each condition. We also recorded the particiapants' computer screens 
throughout the experiment to capture non-interface activity such as 
web searching and inspection of system networking settings.

We ran our experiment at Xlab, the Experimental Social Science Laboratory at University of 
California, Berkeley. The Xlab allows researchers to run experiments on 36 participants at a time.
There are 36 Windows machines in rows of 4, which are separated by cubicle walls. 
We provided the scripts to set up the simulated censorship environment, install necessary software, 
start the video recording, and save the logs and videos. Our experiment was not limited by Xlab
requirements. However, the configuration interface relies on OS-provided features but testing 
was only done on Windows machines, as a byproduct of using Xlab.

@@

Out of our 124 participants, ages ranged from {\color {red} TODO}
(sigma = {\color {red} TODO}). {\color {red} TODO} were male and 
{\color {red} TODO} of our participants had at least a college education.
